Is high social class always beneficial for survival? Northern Sweden 1801--2013

### 2018-03-05 article3d.Rmd
article on oneSpell data

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(engine='R',fig.path='figs/art3-',tidy=FALSE,fig.height=8.5,fig.width=8,strip.white=TRUE,cache=FALSE,echo=FALSE,message=FALSE,comment=NA,show.signif.stars = FALSE)##$
library(knitr)
library(eha)
library(xtable)
library(skum)
options(show.signif.stars = FALSE)
```

```{r readalldata}
source("R/get.file.R")
skel <- get.file("rawskel")
skel$region <- "ske"
skel$urban <- FALSE
vars <- c("id", "birthdate", "sex", "period", "urban", "civst", "hisclass", "enter", "exit", "event")
skel <- skel[, vars]
##
skume <- get.file("rawskume")
skume <- skume[, vars]
##
vb <- rbind(skel, skume)
vb <- vb[order(vb$id, vb$exit), ]
vb$id <- paste("vb", vb$id, sep = "_")
##
foli <- get.file("rawfoli")
foli <- foli[, vars]
```

@@

```{r lexis2,fig.cap="The sampling frame.", fig.height = 5}
plot(c(1801, 1801, 1951, 1951, 1801), c(30, 100, 100, 30, 30),
     type = "l", xlim = c(1800, 2014), lwd = 2, col = "red",
     xlab = "Year", ylab = "Age", axes = FALSE, ylim = c(0, 100))
axis(1, at = c(1801, 1851, 1901, 1951, 1976, 1991, 2014), las = 2)
axis(2, las = 2, at = c(0, 30, 65, 100))
axis(4, las = 2, at = c(0, 30, 65, 100))
box()
abline(h = 0)
lines(c(1976, 1976, 2014, 2014, 1976), c(30, 100, 100, 30, 30),
      col = "blue", lwd = 2)
#
pn.exp <- round(sum(skel$exposure) + sum(skume$exposure))
pn.d <-  as.integer(sum(skel$event) + sum(skume$event))
#
text(1875, 70, "PopLink", cex = 2, col = "red")
##text(1899, 60, paste(pn.rec, "records"))
text(1875, 55, paste(round(pn.exp / 1000000, 1), "million p-years"))
text(1874.5, 50, paste(round(pn.d / 1000, 1), "thousand deaths"))
#
n.exp <- round(sum(foli$exposure))
n.d <-  as.integer(sum(foli$event))

@@

```{r hiscskumm, fig.cap = "Exposure by social class (from the top: elite, middle, workers, farmers, missing) and time period, men.", fig.scap="Exposure by social class and time period, men.", out.height = "14cm", fig.height=4}
xx <- with(skel[skel$sex == "male", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
xy <- with(skume[skume$sex == "male", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
xz <- with(foli[foli$sex == "male", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
xxx <- rbind(xx, xy, xz)
yy <- round(prop.table(xxx, 1) * 100, 1)
barplot(t(yy)[5:1, ], col = c("red", "darkgreen", "blue", "green", "black"), las = 2, ylab = "Per cent")
```

```{r hiscskumw, fig.cap = "Exposure by social class (from the top: elite, middle, workers, missing) and time period, women.", out.height = "14cm", fig.scap="Exposure by social class and time period, women.",fig.height=4}
xx <- with(skel[skel$sex == "female", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
xy <- with(skume[skume$sex == "female", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
xz <- with(foli[foli$sex == "female", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
xxx <- rbind(xx, xy, xz)
yy <- round(prop.table(xxx, 1) * 100, 1)
barplot(t(yy)[5:1, ], col = c("red", "darkgreen", "blue", "green", "black"), las = 2, ylab = "Per cent")
```

@@

```{r civstskum, fig.cap = "Exposure by marital status (from the top: never married, married and previously married) and time period", fig.scap='Exposure by marital status and time period.',out.height = "14cm",fig.height=4}
xx <- with(skel, tapply(exposure, list(period = period, civst = civst), sum))
xy <- with(skume, tapply(exposure, list(period = period, civst = civst), sum))
xz <- with(foli, tapply(exposure, list(period = period, civst = civst), sum))
xxx <- rbind(xx, xy, xz)
yy <- round(prop.table(xxx, 1) * 100, 1)
barplot(t(yy)[3:1, ], col = c("red", "blue", "green"), las = 2, ylab = "Per cent")
```

@@

labb <- levels(vb0Tab$period)
nper <- length(labb)
#source("R/getPHglm.R")
## End 'common'
```

@@

```{r assesciv, fig.cap = "Fraction ever married by hisclass and period.", fig.height = 4}
orr <- aggregate(x = vb0Tab$exposure, by = vb0Tab[, c("period", "civst", "hisclass")], FUN = sum)

levper <- levels(orr$period)
nper <- length(levper)
levhis <- levels(orr$hisclass)
nhis <- length(levhis)
out <- matrix(0, nrow = nhis, ncol = nper)
for (per in 1:nper){
    for (his in 1:nhis){
        oo <- orr[orr$period == levper[per] & orr$hisclass == levhis[his], ]
        nam <- sum(oo$x)
        out[his, per] <- sum(oo$x[oo$civst != "unmarried"]) / nam
    }
}
out <- 100 * out
xli <-  c(1812, 1837, 1862, 1887, 1912, 1932, 1978, 1982, 1988, 1995, 2005, 2011)
plot(xli, out[1, ], type = "b", xlab = "", ylab = "Fraction ever married", 
     axes = FALSE, col = scol[1], lty = 1, pch = 1, ylim = c(0,100))
axis(1, at = xli, labels = levper, las = 2, cex = 0.7)
axis(2, las = 1)
box()
abline(h = 0)
for (r in 2:nhis){
    lines(xli, out[r, ], col = scol[r], lty = r, pch = r, type = "b")
}
legend("bottomright", legend = levhis, lty = 1:5, col = scol, pch = 1:5)
```

@@

```{r assesurb, fig.cap = "Fraction urban by hisclass and period.", fig.height = 4}
orr <- aggregate(x = vb0Tab$exposure, by = vb0Tab[, c("period", "urban", "hisclass")], FUN = sum)

levper <- levels(orr$period)
nper <- length(levper)
levhis <- levels(orr$hisclass)
nhis <- length(levhis)
out <- matrix(0, nrow = nhis, ncol = nper)
for (per in 1:nper){
    for (his in 1:nhis){
        oo <- orr[orr$period == levper[per] & orr$hisclass == levhis[his], ]
        nam <- sum(oo$x)
        out[his, per] <- sum(oo$x[oo$urban == TRUE]) / nam
    }
}
out <- 100 * out
xli <-  c(1812, 1837, 1862, 1887, 1912, 1932, 1978, 1982, 1988, 1995, 2005, 2011)
plot(xli, out[1, ], type = "b", xlab = "", ylab = "Fraction urban", 
     axes = FALSE, col = scol[1], lty = 1, pch = 1, ylim = c(0,100))
axis(1, at = xli, labels = levper, las = 2, cex = 0.7)
axis(2, las = 1)
box()
abline(h = 0)
for (r in 2:nhis){
    lines(xli, out[r, ], col = scol[r], lty = r, pch = r, type = "b")
}
legend("bottomleft", legend = levhis, lty = 1:5, col = scol, pch = 1:5)
```

### Inverse exposure weighting

@@

As an illustration, take the period `1901-1925` with males. There are six combinations of civil status and urban/rural, and for 
each we calculate the total exposure time. The total distribution by confounders is

```{r exweight}
exm <- vb0Tab[vb0Tab$period == "1901-1925" & vb0Tab$sex == "male", c("age", "civst", "urban", "exposure", "event", "hisclass")]
all <- aggregate(exm[, c("exposure", "event")], by = exm[, c("civst", "urban")], FUN = sum)
all
```

and for the elite we have

```{r forelite}
elite <- exm[exm$hisclass == "elite", c("age", "civst", "urban", "exposure", "event", "hisclass")]
el <- aggregate(elite[, c("exposure", "event")], by = elite[, c("civst", "urban")], FUN = sum)
el
```

The weights are now 'all["exposure"]" / el["exposure"]`, that is

```{r wei}
wei <-  all$exposure / el$exposure
all$wei <- wei 
all$Expos <- all$exposure * wei
all$Eve <- all$event * wei
all
```

Put these weights on the data frame `elite`:

```{r putonel}
all$id <- paste(all$civst, all$urban, sep = "_")
elite$id <- paste(elite$civst, elite$urban, sep = "_")
indx <- match(elite$id, all$id)
elite$wght <- all$wei[indx]
```

Now repeat this for the other four hisclasses:

```{r therest}
middle <- exm[exm$hisclass == "middle", c("age", "civst", "urban", "exposure", "event", "hisclass")]
el <- aggregate(middle["exposure"], by = middle[, c("civst", "urban")], FUN = sum)
all$wei <- all$exposure / el$exposure
middle$id <- paste(middle$civst, middle$urban, sep = "_")
indx <- match(middle$id, all$id)
middle$wght <- all$wei[indx]
el
##
worker <- exm[exm$hisclass == "worker", c("age", "civst", "urban", "exposure", "event", "hisclass")]
el <- aggregate(worker["exposure"], by = worker[, c("civst", "urban")], FUN = sum)
all$wei <- all$exposure / el$exposure
worker$id <- paste(worker$civst, worker$urban, sep = "_")
indx <- match(worker$id, all$id)
worker$wght <- all$wei[indx]
el
##
farmer <- exm[exm$hisclass == "farmer", c("age", "civst", "urban", "exposure", "event", "hisclass")]
el <- aggregate(farmer["exposure"], by = farmer[, c("civst", "urban")], FUN = sum)
all$wei <- all$exposure / el$exposure
farmer$id <- paste(farmer$civst, farmer$urban, sep = "_")
indx <- match(farmer$id, all$id)
farmer$wght <- all$wei[indx]
el
##
non <- exm[exm$hisclass == "NA", c("age", "civst", "urban", "exposure", "event", "hisclass")]
el <- aggregate(non["exposure"], by = non[, c("civst", "urban")], FUN = sum)
all$wei <- all$exposure / el$exposure
non$id <- paste(non$civst, non$urban, sep = "_")
indx <- match(non$id, all$id)
non$wght <- all$wei[indx]
el
```

And put together:

```{r puttog}
per1901 <- rbind(elite, middle, worker, farmer, non)
per1901$age <- factor(per1901$age)
#fit <- lm(exposure ~ civst * urban * age, data = per1901)
#pred <- predict(fit, type = "response")
#per1901$wght2 <- pred / per1901$exposure
```

Let us fit with and without weights. Without weights, see Figure \@ref(fig:testfit).

```{r testfit, fig.cap = "Survival by hisclass, men 1901-1925, non-weighted analysis.", fig.height = 4}
fit1 <- glm(event ~ offset(log(exposure)) + age, data = per1901[per1901$hisclass == "elite", ], family = poisson)
co1 <- fit1$coefficients
co1[2:length(co1)] <- co1[2:length(co1)] + co1[1]
co1 <- exp(co1)
##
fit2 <- glm(event ~ offset(log(exposure)) + age, data = per1901[per1901$hisclass == "middle", ], family = "poisson")
co2 <- fit2$coefficients
co2[2:length(co2)] <- co2[2:length(co2)] + co2[1]
co2 <- exp(co2)
##
fit3 <- glm(event ~ offset(log(exposure)) + age, data = per1901[per1901$hisclass == "worker", ], family = "poisson")
co3 <- fit3$coefficients
co3[2:length(co3)] <- co3[2:length(co3)] + co3[1]
co3 <- exp(co3)
##
fit4 <- glm(event ~ offset(log(exposure)) + age, data = per1901[per1901$hisclass == "farmer", ], family = "poisson")
co4 <- fit4$coefficients
co4[2:length(co4)] <- co4[2:length(co4)] + co4[1]
co4 <- exp(co4)
##
fit5 <- glm(event ~ offset(log(exposure)) + age, data = per1901[per1901$hisclass == "NA", ], family = "poisson")
co5 <- fit5$coefficients
co5[2:length(co5)] <- co5[2:length(co5)] + co5[1]
co5 <- exp(co5)
###
cuts <- seq(5, 55, by = 5)
x <- seq(0, 60, length = 1000)
s1 <- ppch(x, cuts = cuts, levels = co1[1:12], lower.tail = FALSE)
plot(x, s1, type = "l", col = scol[1], axes = FALSE, xlab = "age", ylab = "Surviving fraction after 30.")
axis(1, at = seq(0, 60, by = 10), labels = seq(30, 90, by = 10))
box()
axis(2, las = 1)
abline(h = 0)
abline(v = 0)
s2 <- ppch(x, cuts = cuts, levels = co2[1:12], lower.tail = FALSE)
lines(x, s2, col = scol[2], lty = 2)
s3 <- ppch(x, cuts = cuts, levels = co3[1:12], lower.tail = FALSE)
lines(x, s3, col = scol[3], lty = 3)
s4 <- ppch(x, cuts = cuts, levels = co4[1:12], lower.tail = FALSE)
lines(x, s4, col = scol[4], lty = 4)
s5 <- ppch(x, cuts = cuts, levels = co5[1:12], lower.tail = FALSE)
lines(x, s5, col = scol[5], lty = 5)
##
```

With weights, see Figure \@ref(fig:testfit2):

```{r testfit2, fig.cap = "Survival by hisclass, men 1901-1925, weighted analysis.", fig.height = 4}
fit1 <- glm(event ~ offset(log(exposure)) + age, data = per1901[per1901$hisclass == "elite", ], family = poisson, weights = wght)
co1 <- fit1$coefficients
co1[2:length(co1)] <- co1[2:length(co1)] + co1[1]
co1 <- exp(co1)
##
fit2 <- glm(event ~ offset(log(exposure)) + age, data = per1901[per1901$hisclass == "middle", ], family = "poisson", weights = wght)
co2 <- fit2$coefficients
co2[2:length(co2)] <- co2[2:length(co2)] + co2[1]
co2 <- exp(co2)
##
fit3 <- glm(event ~ offset(log(exposure)) + age, data = per1901[per1901$hisclass == "worker", ], family = "poisson", weights = wght)
co3 <- fit3$coefficients
co3[2:length(co3)] <- co3[2:length(co3)] + co3[1]
co3 <- exp(co3)
##
fit4 <- glm(event ~ offset(log(exposure)) + age, data = per1901[per1901$hisclass == "farmer", ], family = "poisson", weights = wght)
co4 <- fit4$coefficients
co4[2:length(co4)] <- co4[2:length(co4)] + co4[1]
co4 <- exp(co4)
##
fit5 <- glm(event ~ offset(log(exposure)) + age, data = per1901[per1901$hisclass == "NA", ], family = "poisson", weights = wght)
co5 <- fit5$coefficients
co5[2:length(co5)] <- co5[2:length(co5)] + co5[1]
co5 <- exp(co5)
###
cuts <- seq(5, 55, by = 5)
x <- seq(0, 60, length = 1000)
s1 <- ppch(x, cuts = cuts, levels = co1[1:12], lower.tail = FALSE)
plot(x, s1, type = "l", col = scol[1], axes = FALSE, xlab = "Age", ylab = "Surviving fraction after 30")
axis(1, at = seq(0, 60, by = 10), labels = seq(30, 90, by = 10))
axis(2, las = 1)
box()
abline(h = 0)
abline(v = 0)
s2 <- ppch(x, cuts = cuts, levels = co2[1:12], lower.tail = FALSE)
lines(x, s2, col = scol[2], lty = 2)
s3 <- ppch(x, cuts = cuts, levels = co3[1:12], lower.tail = FALSE)
lines(x, s3, col = scol[3], lty = 3)
s4 <- ppch(x, cuts = cuts, levels = co4[1:12], lower.tail = FALSE)
lines(x, s4, col = scol[4], lty = 4)
s5 <- ppch(x, cuts = cuts, levels = co5[1:12], lower.tail = FALSE)
lines(x, s5, col = scol[5], lty = 5)
##
```

>>>>

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(engine='R',fig.path='figs/art3d-',tidy=FALSE,fig.height=8.5,fig.width=8,strip.white=TRUE,cache=FALSE,echo=FALSE,message=FALSE,comment=NA,show.signif.stars = FALSE)##$
library(knitr)
library(eha)
library(xtable)
library(skum)
source("R/oneSpell.R")
options(show.signif.stars = FALSE)
```

```{r readalldata}
load("data/skume.rda")
skume <- skume[order(skume$id, skume$enter), ]
skume$id <- paste("pop", skume$id, sep = "_")
##
skume1 <- age.window(skume, c(40, 65))
skume1 <- oneSpell(skume1)
##
skume2 <- age.window(skume, c(65, 100))
skume2 <- oneSpell(skume2)
##################################################
load("data/foli.rda")
foli <- foli[order(foli$id, foli$enter), ]
foli$AR <- foli$ULORSAK <- NULL
##
foli1 <- age.window(foli, c(40, 65))
foli1 <- oneSpell(foli1)
##
foli2 <- age.window(foli, c(65, 100))
foli2 <- oneSpell(foli2)
##################################################
source("CreData/R/addPer.R")
skume1 <- addPer(skume1, cuts = seq(1801, 1951, by = 25))
skume2 <- addPer(skume2, cuts = seq(1801, 1951, by = 25))
##
foli1 <- addPer(foli1, cuts = c(1976, 1986, 1996, 2005, 2014))
foli2 <- addPer(foli2, cuts = c(1976, 1986, 1996, 2005, 2014))
##################################################
vb1 <- rbind(skume1, foli1)
vb1$period <- factor(vb1$period)
##
vb2 <- rbind(skume2, foli2)
vb2$period <- factor(vb2$period)
vb <- rbind(vb1, vb2)
vb$exposure <- vb$exit - vb$enter # For statistics and figures purposes
```

@@

```{r lexis2,fig.cap="The sampling frame.", fig.height = 5}
plot(c(1801, 1801, 1951, 1951, 1801), c(30, 100, 100, 30, 30),
     type = "l", xlim = c(1800, 2014), lwd = 2, col = "red",
     xlab = "Year", ylab = "Age", axes = FALSE, ylim = c(0, 100))
axis(1, at = c(1801, 1851, 1901, 1951, 1976, 1986, 1996, 2005, 2014), las = 2)
axis(2, las = 2, at = c(0, 30, 65, 100))
axis(4, las = 2, at = c(0, 30, 65, 100))
box()
abline(h = 0)
lines(c(1976, 1976, 2014, 2014, 1976), c(30, 100, 100, 30, 30),
      col = "blue", lwd = 2)
#
pn.exp <- round(sum(skume1$exit - skume1$enter) + sum(skume2$exit - skume2$enter))
pn.d <-  as.integer(sum(skume1$event) + sum(skume2$event))
#
text(1875, 70, "PopLink", cex = 2, col = "red")
##text(1899, 60, paste(pn.rec, "records"))
text(1875, 55, paste(round(pn.exp / 1000000, 1), "million p-years"))
text(1874.5, 50, paste(round(pn.d / 1000, 1), "thousand deaths"))
#
n.exp <- round(sum(foli1$exit - foli1$enter) + sum(foli2$exit -foli2$enter))
n.d <-  as.integer(sum(foli1$event) + sum(foli2$event))

@@

```{r hiscskumm, fig.cap = "Exposure by social class (from the top: elite, middle, workers, farmers, none) and time period, men.", fig.scap="Exposure by social class and time period, men.", out.height = "14cm", fig.height=4}
xx <- with(vb[vb$sex == "male", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
##xy <- with(skume[skume$sex == "male", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
##xz <- with(foli[foli$sex == "male", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
##xxx <- rbind(xx, xy, xz)
yy <- round(prop.table(xx, 1) * 100, 1)
barplot(t(yy)[5:1, ], col = c("red", "darkgreen", "blue", "green", "black"), las = 2, ylab = "Per cent")
```

```{r hiscskumw, fig.cap = "Exposure by social class (from the top: elite, middle, workers, missing) and time period, women.", out.height = "14cm", fig.scap="Exposure by social class and time period, women.",fig.height=4}
xx <- with(vb[vb$sex == "female", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
##xy <- with(skume[skume$sex == "female", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
##xz <- with(foli[foli$sex == "female", ], tapply(exposure, list(period = period, hisclass = hisclass), sum))
##xxx <- rbind(xx, xy, xz)
yy <- round(prop.table(xx, 1) * 100, 1)
barplot(t(yy)[5:1, ], col = c("red", "darkgreen", "blue", "green", "black"), las = 2, ylab = "Per cent")
```

@@

```{r civstskum, fig.cap = "Exposure by marital status (from the top: never married, married and previously married) and time period", fig.scap='Exposure by marital status and time period.',out.height = "14cm",fig.height=4}
xx <- with(vb, tapply(exposure, list(period = period, civst = civst), sum))
##xy <- with(skume, tapply(exposure, list(period = period, civst = civst), sum))
##xz <- with(foli, tapply(exposure, list(period = period, civst = civst), sum))
##xxx <- rbind(xx, xy, xz)
yy <- round(prop.table(xx, 1) * 100, 1)
barplot(t(yy)[3:1, ], col = c("red", "blue", "green"), las = 2, ylab = "Per cent")
```

@@

labb <- levels(vb$period)
nper <- length(labb)
#source("R/getPHglm.R")
## End 'common'
```

@@

```{r assesciv, fig.cap = "Fraction ever married by hisclass and period.", fig.height = 4}
orr <- aggregate(x = vb$exposure, by = vb[, c("period", "civst", "hisclass")], FUN = sum)

levper <- levels(orr$period)
nper <- length(levper)
levhis <- levels(orr$hisclass)
nhis <- length(levhis)
out <- matrix(0, nrow = nhis, ncol = nper)
for (per in 1:nper){
    for (his in 1:nhis){
        oo <- orr[orr$period == levper[per] & orr$hisclass == levhis[his], ]
        nam <- sum(oo$x)
        out[his, per] <- sum(oo$x[oo$civst != "unmarried"]) / nam
    }
}
out <- 100 * out
xli <-  c(1812, 1837, 1862, 1887, 1912, 1932, 1981, 1991, 2000, 2009)
plot(xli, out[1, ], type = "b", xlab = "", ylab = "Fraction ever married", 
     axes = FALSE, col = scol[1], lty = 1, pch = 1, ylim = c(0,100))
axis(1, at = xli, labels = levper, las = 2, cex = 0.7)
axis(2, las = 1)
box()
abline(h = 0)
for (r in 2:nhis){
    lines(xli, out[r, ], col = scol[r], lty = r, pch = r, type = "b")
}
legend("bottomright", legend = levhis, lty = 1:5, col = scol, pch = 1:5)
```

@@

```{r assesurb, fig.cap = "Fraction urban by hisclass and period.", fig.height = 4}
orr <- aggregate(x = vb$exposure, by = vb[, c("period", "urban", "hisclass")], FUN = sum)

levper <- levels(orr$period)
nper <- length(levper)
levhis <- levels(orr$hisclass)
nhis <- length(levhis)
out <- matrix(0, nrow = nhis, ncol = nper)
for (per in 1:nper){
    for (his in 1:nhis){
        oo <- orr[orr$period == levper[per] & orr$hisclass == levhis[his], ]
        nam <- sum(oo$x)
        out[his, per] <- sum(oo$x[oo$urban == TRUE]) / nam
    }
}
out <- 100 * out
xli <-  c(1812, 1837, 1862, 1887, 1912, 1932, 1981, 1991, 2000, 2009)
plot(xli, out[1, ], type = "b", xlab = "", ylab = "Fraction urban", 
     axes = FALSE, col = scol[1], lty = 1, pch = 1, ylim = c(0,100))
axis(1, at = xli, labels = levper, las = 2, cex = 0.7)
axis(2, las = 1)
box()
abline(h = 0)
for (r in 2:nhis){
    lines(xli, out[r, ], col = scol[r], lty = r, pch = r, type = "b")
}
legend("bottomleft", legend = levhis, lty = 1:5, col = scol, pch = 1:5)
```

### Inverse probability weighting

@@

As an illustration, take the period `2005-2013` with males, 65--99 years of age. There are six combinations of civil status and urban/rural, and for 
each we calculate the total exposure time. The total distribution by confounders is

```{r exweight, echo = TRUE}
vb2_2005_men <- vb2[vb2$sex == "male" & vb2$period == "2005-2013", ]
fit1 <- glm(I(hisclass == "elite") ~ urban * civst, data = vb2_2005_men, family = binomial)
fit2 <- glm(I(hisclass == "middle") ~ urban * civst, data = vb2_2005_men, family = binomial)
fit3 <- glm(I(hisclass == "worker") ~ urban * civst, data = vb2_2005_men, family = binomial)
fit4 <- glm(I(hisclass == "farmer") ~ urban * civst, data = vb2_2005_men, family = binomial)
fit5 <- glm(I(hisclass == "none") ~ urban * civst, data = vb2_2005_men, family = binomial)
wght <- numeric(NROW(vb2_2005_men))
##
gg <- vb2_2005_men$hisclass == "elite"
wght[gg] <- 1 / predict(fit1, type = "response")[gg]
##
gg <- vb2_2005_men$hisclass == "middle"
wght[gg] <- 1 / predict(fit1, type = "response")[gg]
##
gg <- vb2_2005_men$hisclass == "worker"
wght[gg] <- 1 / predict(fit1, type = "response")[gg]
##
gg <- vb2_2005_men$hisclass == "farmer"
wght[gg] <- 1 / predict(fit1, type = "response")[gg]
##
gg <- vb2_2005_men$hisclass == "none"
wght[gg] <- 1 / predict(fit1, type = "response")[gg]
##
vb2_2005_men$wght <- wght
```

Let us fit with and without weights, see Figure \@ref(fig:testfit).

```{r testfit, fig.cap = "Survival by hisclass, men 1901-1925, non-weighted analysis.", fig.height = 4}
fit <- coxph(Surv(enter, exit, event) ~ strata(hisclass), data = vb2_2005_men)
fitw <- coxph(Surv(enter, exit, event) ~ strata(hisclass), weights = wght, data = vb2_2005_men)
oldpar <- par(mfrow = c(1, 2))
plot(survfit(fit), mark.time = FALSE, col = scol, lty = 1:5, xlim = c(65, 100), xlab = "Age", ylab = "Survival", main = "Unweighted")
abline(h = 0)
abline(h = 0.5, lty = 3)
med <- summary(survfit(fit))$table[, "median"]
lines(c(med["none"], med["none"]), c(0, 0.5), col = "red", lty = 3)
lines(c(med["elite"], med["elite"]), c(0, 0.5), col = "black", lty = 3)
plot(survfit(fitw), mark.time = FALSE, col = scol, lty = 1:5, xlim = c(65, 100), xlab = "Age", ylab = "Survival", main = "Weighted")
abline(h = 0)
abline(h = 0.5, lty = 3)
medw <- summary(survfit(fitw))$table[, "median"]
lines(c(medw["none"], medw["none"]), c(0, 0.5), col = "red", lty = 3)
lines(c(medw["elite"], medw["elite"]), c(0, 0.5), col = "black", lty = 3)
medi <- cbind(med, medw)
colnames(medi) <- c("Unweighted", "weighted")
##
```

The estimated medians are

```{r estmed}
knitr::kable(round(medi, 1))
```
